---
title: "AFDM vs Salinity GLMM - Clean, Publication-Ready"
output: html_notebook
---

## Data Analysis File for Organic Matter manuscript by Tabor et al.

# Modeling the organic matter - SC relationships

1.  Load libraries

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
```

2.  Read data

```{r}
setwd("~/Downloads/Updated code + summary sheets")

df <- read.csv("SALTY_C_AFDM_YEAR_1_ALL.csv", check.names=FALSE) %>%
  mutate(
    Site  = factor(Site),
    OM    = factor(OM),  # e.g., CBOM, FBOM, Periphyton
    month = factor(Season, levels=c("February","May","August","October")),
    SC    = as.numeric(SC),
    AFDM  = as.numeric(AFDM)
  )
```

3.  Plot aesthetics

```{r}
OM_colors <- c("CBOM"="#E69F00", "FBOM"="#0072B2", "Periphyton"="#228B22")
OM_shapes <- c("CBOM"=16, "FBOM"=17, "Periphyton"=15)

month_labels <- c(
  "February" = "February (Winter 2024)",
  "May"      = "May (Spring 2024)",
  "August"   = "August (Summer 2024)",
  "October"  = "October (Autumn 2023)"
)

SC_seq <- seq(min(df$SC, na.rm=TRUE), max(df$SC, na.rm=TRUE), length.out=100)
```

4.  Fit full model

```{r}
mod_AFDM <- lmer(AFDM ~ SC * month * OM + (1 | Site), data=df)
```

5.  Extract all SC slopes (table)

```{r}
AFDM_SC_slopes_all <- emtrends(mod_AFDM, ~ month | OM, var="SC") %>%
  summary(infer=TRUE) %>%
  as.data.frame() %>%
  mutate(
    t_value = SC.trend / SE,
    p_value = 2 * pt(-abs(t_value), df),
    Significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ "ns"
    ),
    SC_slope  = round(SC.trend, 2),
    Std_Error = round(SE, 2),
    t_value   = round(t_value, 2),
    p_value   = signif(p_value, 3),
    lower.CL  = round(lower.CL, 2),
    upper.CL  = round(upper.CL, 2)
  ) %>%
  select(OM, month, SC_slope, Std_Error, df, t_value, p_value, lower.CL, upper.CL, Significance)

# Print full slopes table
print("=== AFDM ~ SC Slopes by OM and Month ===")
print(AFDM_SC_slopes_all)

# Save table for manuscript
write.csv(AFDM_SC_slopes_all, "~/Downloads/Updated code + summary sheets/AFDM_SC_slopes_by_OM_and_Month.csv", row.names=FALSE)
```

6.  Extract only significant slopes for plotting

```{r}
AFDM_SC_slopes_sig <- AFDM_SC_slopes_all %>% filter(Significance != "ns")

pred_sig <- emmeans(mod_AFDM, ~ SC | month * OM, at=list(SC=SC_seq), cov.reduce=FALSE) %>%
  as.data.frame() %>%
  semi_join(AFDM_SC_slopes_sig, by=c("month","OM"))
```

7.  Plot significant slopes

```{r}
p_AFDM_sig <- ggplot(df, aes(SC, AFDM, color=OM, shape=OM)) +
  geom_jitter(width=0.1, size=3, alpha=0.7) +
  geom_ribbon(data=pred_sig,
              aes(x=SC, ymin=lower.CL, ymax=upper.CL, group=interaction(month,OM)),
              inherit.aes=FALSE, fill="grey70", alpha=0.3) +
  geom_line(data=pred_sig,
            aes(x=SC, y=emmean, group=interaction(month,OM)),
            inherit.aes=FALSE, color="black", size=1.2) +
  facet_grid(OM ~ month, scales="free_y", labeller=labeller(month=month_labels)) +
  scale_color_manual(values=OM_colors) +
  scale_shape_manual(values=OM_shapes) +
  theme_bw(base_size=16) +
  theme(strip.background=element_blank(),
        strip.text=element_text(size=14, face="bold")) +
  labs(x="Specific Conductivity (µS/m)", y="AFDM (g/m²)")

# Print figure
print(p_AFDM_sig)
```

# Chl-a - AC relationsihps (periphyton only)

1.  Read and clean data

```{r}

df <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_CHLA_ROCKS.csv", check.names=FALSE) %>%
  mutate(
    Site  = factor(Site),
    month = factor(Season, levels=c("February","May","August","October")),
    SC    = as.numeric(SC),
    CHLA  = as.numeric(CHLA)
  )
```

2.  Prediction sequence

```{r}
SC_seq <- seq(min(df$SC, na.rm=TRUE), max(df$SC, na.rm=TRUE), length.out=100)
```

3.  Fit GLMM

```{r}
mod_CHLA <- lmer(CHLA ~ SC * month + (1 | Site), data=df)
```

4.  Extract SC slopes from model

```{r}
CHLA_SC_slopes_all <- emtrends(mod_CHLA, ~ month, var="SC") %>%
  summary(infer=TRUE) %>%
  as.data.frame() %>%
  mutate(
    SC_slope  = signif(SC.trend, 3),
    Std_Error = signif(SE, 3),
    t_value   = signif(t.ratio, 3),
    p_value   = signif(p.value, 3),
    lower.CL  = signif(lower.CL, 3),
    upper.CL  = signif(upper.CL, 3),
    Significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ "ns"
    )
  ) %>%
  select(month, SC_slope, Std_Error, df, t_value, p_value, lower.CL, upper.CL, Significance)

# Print full slopes table
print("=== CHLA ~ SC Slopes by Month ===")
print(CHLA_SC_slopes_all)

# Save table for manuscript
write.csv(CHLA_SC_slopes_all, "~/Downloads/Updated code + summary sheets/CHLA_SC_slopes_by_month.csv", row.names=FALSE)
```

5.  Extract only significant slopes for plotting

```{r}
CHLA_SC_slopes_sig <- CHLA_SC_slopes_all %>% filter(Significance != "ns") %>% select(month)

pred_sig <- emmeans(mod_CHLA, ~ SC | month, at=list(SC=SC_seq), cov.reduce=FALSE) %>%
  as.data.frame() %>%
  semi_join(CHLA_SC_slopes_sig, by="month")
```

6.  Plot significant slopes

```{r}
algae_color <- "#228B22"  # Periphyton

month_labels <- c(
  "February" = "February (Winter 2024)",
  "May"      = "May (Spring 2024)",
  "August"   = "August (Summer 2024)",
  "October"  = "October (Autumn 2024)"
)

p_CHLA_sig <- ggplot(df, aes(x=SC, y=CHLA)) +
  geom_jitter(color=algae_color, size=3, width=0.1, alpha=0.7) +
  geom_ribbon(data=pred_sig,
              aes(x=SC, ymin=lower.CL, ymax=upper.CL),
              fill=algae_color, alpha=0.3, inherit.aes=FALSE) +
  geom_line(data=pred_sig, aes(x=SC, y=emmean),
            color=algae_color, size=1.2, inherit.aes=FALSE) +
  facet_wrap(~ month, scales="fixed", labeller=labeller(month=month_labels)) +
  theme_bw(base_size=16) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size=14, face="bold")
  ) +
  labs(
    x = "Specific Conductivity (µS/m)",
    y = expression("Periphyton chlorophyll " * italic(a) * " concentration (g/"*m^2*")")
  )

# Print figure
print(p_CHLA_sig)
```

# Analysising the isotopes

1.  Read the data

```{r}
df <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_SI_YEAR_1_ALL.csv", check.names = FALSE)
```

2.  Data preparation

```{r}
df <- df %>%
  mutate(
    Season = factor(Season, levels = c("February", "May", "August", "October")),
    Site   = factor(Site),
    OM     = factor(OM),
    SC     = as.numeric(SC),
    d13C   = as.numeric(d13C),
    d15N   = as.numeric(d15N)
  )
```

3.  Fit full interaction models

```{r}
mod_d13C <- lmer(d13C ~ SC * Season * OM + (1 | Site), data = df)
mod_d15N <- lmer(d15N ~ SC * Season * OM + (1 | Site), data = df)
```

4.  Function to extract SC slopes for each Season × OM

```{r}
extract_slopes_all <- function(model, isotope_name) {
  emtrends(model, ~ Season | OM, var = "SC") %>%
    summary(infer = TRUE) %>%
    as.data.frame() %>%
    mutate(
      t_value = SC.trend / SE,
      p_value = 2 * pt(-abs(t_value), df),
      Significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01  ~ "**",
        p_value < 0.05  ~ "*",
        TRUE            ~ "ns"
      ),
      Isotope   = isotope_name,
      SC_slope  = signif(SC.trend, 3),
      Std_Error = signif(SE, 3),
      t_value   = signif(t_value, 3),
      p_value   = signif(p_value, 3)
    ) %>%
    select(Isotope, OM, Season, SC_slope, Std_Error, df, t_value, p_value, Significance)
}
```

5.  Extract slopes for δ13C and δ15N

```{r}
d13C_SC_slopes <- extract_slopes_all(mod_d13C, "d13C")
d15N_SC_slopes <- extract_slopes_all(mod_d15N, "d15N")
```

6.  Print tables separately

```{r}
cat("\n=== δ13C SC Slopes (All) ===\n")
print(d13C_SC_slopes)

cat("\n=== δ15N SC Slopes (All) ===\n")
print(d15N_SC_slopes)
```

7.  Salinity prediction sequence

```{r}
SC_seq <- seq(min(df$SC, na.rm = TRUE), max(df$SC, na.rm = TRUE), length.out = 100)
```

8.  Predicted values by Season × OM

```{r}
emm_d13C <- emmeans(mod_d13C, ~ SC | Season + OM, at = list(SC = SC_seq))
emm_d15N <- emmeans(mod_d15N, ~ SC | Season + OM, at = list(SC = SC_seq))

pred_fixed <- bind_rows(
  as.data.frame(emm_d13C) %>% mutate(Isotope = "d13C"),
  as.data.frame(emm_d15N) %>% mutate(Isotope = "d15N")
)

# Keep only significant Season × OM slopes for plotting
pred_fixed_sig <- pred_fixed %>%
  semi_join(bind_rows(
    d13C_SC_slopes %>% filter(Significance != "ns"),
    d15N_SC_slopes %>% filter(Significance != "ns")
  ), by = c("Season", "OM", "Isotope"))
```

9.  Pivot raw data to long format

```{r}
df_long <- df %>%
  pivot_longer(cols = c(d13C, d15N), names_to = "Isotope", values_to = "Value")
```

10. Plot aesthetics

```{r}
OM_shapes <- c("CBOM" = 16, "FBOM" = 17, "Periphyton" = 15)
OM_colors <- c("CBOM" = "#E69F00", "FBOM" = "#0072B2", "Periphyton" = "#228B22")
month_labels <- c(
  "February" = "February (Winter 2024)",
  "May"      = "May (Spring 2024)",
  "August"   = "August (Summer 2024)",
  "October"  = "October (Autumn 2023)"
)
```

11. δ13C Plot

```{r}
p_d13C <- ggplot(df_long %>% filter(Isotope == "d13C"),
                 aes(x = SC, y = Value, color = OM, shape = OM)) +
  geom_jitter(size = 3, width = 0.1, alpha = 0.7) +
  geom_ribbon(data = pred_fixed_sig %>% filter(Isotope == "d13C"),
              aes(x = SC, ymin = lower.CL, ymax = upper.CL,
                  fill = OM, group = interaction(Season, OM)),
              alpha = 0.3, inherit.aes = FALSE) +
  geom_line(data = pred_fixed_sig %>% filter(Isotope == "d13C"),
            aes(x = SC, y = emmean, color = OM, group = interaction(Season, OM)),
            size = 1.2, inherit.aes = FALSE) +
  scale_color_manual(values = OM_colors, guide = "legend") +
  scale_fill_manual(values = OM_colors, guide = "none") +
  scale_shape_manual(values = OM_shapes) +
  facet_wrap(~ Season, scales = "free_y", labeller = labeller(Season = month_labels)) +
  theme_bw(base_size = 16) +
  labs(x = "Specific Conductivity (µS/m)",
       y = expression(delta^{13} * "C"),
) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))

print(p_d13C)
```

12. δ15N Plot

```{r}
p_d15N <- ggplot(df_long %>% filter(Isotope == "d15N"),
                 aes(x = SC, y = Value, color = OM, shape = OM)) +
  geom_jitter(size = 3, width = 0.1, alpha = 0.7) +
  geom_ribbon(data = pred_fixed_sig %>% filter(Isotope == "d15N"),
              aes(x = SC, ymin = lower.CL, ymax = upper.CL,
                  fill = OM, group = interaction(Season, OM)),
              alpha = 0.3, inherit.aes = FALSE) +
  geom_line(data = pred_fixed_sig %>% filter(Isotope == "d15N"),
            aes(x = SC, y = emmean, color = OM, group = interaction(Season, OM)),
            size = 1.2, inherit.aes = FALSE) +
  scale_color_manual(values = OM_colors, guide = "legend") +
  scale_fill_manual(values = OM_colors, guide = "none") +
  scale_shape_manual(values = OM_shapes) +
  facet_wrap(~ Season, scales = "free_y", labeller = labeller(Season = month_labels)) +
  theme_bw(base_size = 16) +
  labs(x = "Specific Conductivity (µS/m)",
       y = expression(delta^{15} * "N"),
) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
print(p_d15N)
```

# C:N ratios

1.  Read data

```{r}
df <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_CN_YEAR_1_ALL.csv", check.names = FALSE)
```

2.  Data preparation

```{r}
df <- df %>%
  mutate(
    Site = factor(Site),
    month = factor(Season, levels = c("February", "May", "August", "October")),
    OM = factor(OM),
    SC = as.numeric(SC),
    C_N = as.numeric(C_N)
  )
```

3.  Fit Full GLMM

```{r}
mod_CN <- lmer(
  C_N ~ SC * month * OM + (1 | Site),
  data = df
)
```

4.  Extract SC slopes for each Month × OM

```{r}
CN_SC_slopes <- emtrends(
  mod_CN,
  ~ month | OM,
  var = "SC"
) %>%
  summary(infer = TRUE) %>%
  as.data.frame() %>%
  mutate(
    t_value = SC.trend / SE,
    p_value = 2 * pt(-abs(t_value), df),
    Significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ "ns"
    ),
    SC_slope = signif(SC.trend, 3),
    Std_Error = signif(SE, 3),
    t_value = signif(t_value, 3),
    p_value = signif(p_value, 3),
    lower.CL = signif(lower.CL, 3),
    upper.CL = signif(upper.CL, 3)
  ) %>%
  select(OM, month, SC_slope, Std_Error, df, t_value, p_value, lower.CL, upper.CL, Significance)
```

5.  Print SC slopes table (all)

```{r}
print("=== C:N ~ SC Slopes by Month × OM (All Slopes) ===")
print(CN_SC_slopes)
write.csv(CN_SC_slopes, "~/Downloads/Updated code + summary sheets/CN_SC_slopes_by_month_OM.csv", row.names = FALSE)
```

6.  Salinity prediction sequence

```{r}
SC_seq <- seq(
  min(df$SC, na.rm = TRUE),
  max(df$SC, na.rm = TRUE),
  length.out = 100
)
```

7.  Predicted values for Month × OM

```{r}
emm_CN <- emmeans(
  mod_CN,
  ~ SC | month + OM,
  at = list(SC = SC_seq)
)
pred_CN <- as.data.frame(emm_CN)
```

8.  Keep only significant slopes for plotting

```{r}
sig_CN_slopes <- CN_SC_slopes %>%
  filter(Significance != "ns") %>%
  select(OM, month)

pred_CN_sig <- pred_CN %>%
  semi_join(sig_CN_slopes, by = c("OM", "month"))
```

9.  Plot settings

```{r}
OM_shapes <- c("CBOM" = 16, "FBOM" = 17, "Periphyton" = 15)
OM_colors <- c("CBOM" = "#E69F00",
               "FBOM" = "#0072B2",
               "Periphyton" = "#228B22")

month_labels <- c(
  "February" = "February (Winter 2024)",
  "May"      = "May (Spring 2024)",
  "August"   = "August (Summer 2024)",
  "October"  = "October (Autumn 2024)"
)
```

10. Faceted Plot: OM × Month (only significant lines)

```{r}
p_CN <- ggplot(df, aes(x = SC, y = C_N, color = OM, shape = OM)) +
  
  # Raw data points
  geom_jitter(size = 3, width = 0.1, alpha = 0.7) +
  
  # Only significant ribbons
  geom_ribbon(
    data = pred_CN_sig,
    aes(x = SC,
        ymin = lower.CL,
        ymax = upper.CL,
        fill = OM,
        group = interaction(OM, month)),
    alpha = 0.3,
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  
  # Only significant lines
  geom_line(
    data = pred_CN_sig,
    aes(x = SC,
        y = emmean,
        color = OM,
        group = interaction(OM, month)),
    size = 1.2,
    inherit.aes = FALSE
  ) +
  
  # Manual scales
  scale_color_manual(values = OM_colors) +
  scale_fill_manual(values = OM_colors) +
  scale_shape_manual(values = OM_shapes) +
  
  # Facet grid: OM rows × Month columns
  facet_grid(
    OM ~ month,
    labeller = labeller(month = month_labels),
    scales = "free_y"
  ) +
  
  # Theme + labels
  theme_bw(base_size = 16) +
  labs(
    x = "Specific Conductivity (µS/m)",
    y = "C:N"
  ) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 14, face = "bold")
  )

# Print the faceted plot
print(p_CN)
```

# δ13C vs δ15N Biplot

1.  Load data

```{r}
df <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_SI_YEAR_1_ALL.csv", check.names = FALSE)
```

2.  Prepare data

```{r}
df <- df %>%
  mutate(
    Season = factor(Season, levels = c("February", "May", "August", "October")),
    Site   = factor(Site),
    OM     = factor(OM),
    SC     = as.numeric(SC),
    d13C   = as.numeric(d13C),
    d15N   = as.numeric(d15N)
  )
```

3.  Fit full interaction models

```{r}
mod_d13C <- lmer(d13C ~ SC * Season * OM + (1 | Site), data = df)
mod_d15N <- lmer(d15N ~ SC * Season * OM + (1 | Site), data = df)
```

4.  Extract SC slopes and significance

```{r}
extract_slopes <- function(model, isotope) {
  emtrends(model, ~ Season | OM, var = "SC") %>%
    summary(infer = TRUE) %>%
    as.data.frame() %>%
    mutate(
      t_value = SC.trend / SE,
      p_value = 2 * pt(-abs(t_value), df),
      Significance = case_when(
        p_value < 0.001 ~ "***",
        p_value < 0.01  ~ "**",
        p_value < 0.05  ~ "*",
        TRUE            ~ "ns"
      ),
      Isotope = isotope,
      SC_slope = SC.trend
    ) %>%
    select(Isotope, OM, Season, SC_slope, SE, df, t_value, p_value, Significance)
}

slopes_d13C <- extract_slopes(mod_d13C, "d13C")
slopes_d15N <- extract_slopes(mod_d15N, "d15N")

# Combine slope tables
all_slopes <- bind_rows(slopes_d13C, slopes_d15N)
cat("\n=== SC Slopes for δ13C and δ15N ===\n")
print(all_slopes)
```

5.  Salinity sequence for predictions

```{r}
SC_seq <- seq(min(df$SC, na.rm = TRUE), max(df$SC, na.rm = TRUE), length.out = 100)
```

6.  Predicted values (only significant slopes)

```{r}
emm_d13C <- emmeans(mod_d13C, ~ SC | Season + OM, at = list(SC = SC_seq))
emm_d15N <- emmeans(mod_d15N, ~ SC | Season + OM, at = list(SC = SC_seq))

pred <- bind_rows(
  as.data.frame(emm_d13C) %>% mutate(Isotope = "d13C"),
  as.data.frame(emm_d15N) %>% mutate(Isotope = "d15N")
)

sig_slopes <- bind_rows(
  slopes_d13C %>% filter(Significance != "ns"),
  slopes_d15N %>% filter(Significance != "ns")
)

pred_sig <- pred %>% semi_join(sig_slopes, by = c("Season", "OM", "Isotope"))
```

7.  Pivot predictions to wide for biplot

```{r}
pred_wide <- pred_sig %>%
  pivot_wider(names_from = Isotope, values_from = emmean) 
```

8.  Plot aesthetics

```{r}
OM_shapes <- c("CBOM" = 16, "FBOM" = 17, "Periphyton" = 15)
OM_colors <- c("CBOM" = "#E69F00", "FBOM" = "#0072B2", "Periphyton" = "#228B22")
month_labels <- c(
  "February" = "February (Winter 2024)",
  "May"      = "May (Spring 2024)",
  "August"   = "August (Summer 2024)",
  "October"  = "October (Autumn 2024)"
)
```

9.  Biplot δ13C vs δ15N with significant slopes

```{r}
ggplot(df, aes(x = d13C, y = d15N, color = OM, shape = OM)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(data = pred_wide, aes(x = d13C, y = d15N, group = interaction(OM, Season)),
            size = 1.2, inherit.aes = FALSE) +
  geom_vline(xintercept= -28, color = "darkred", linetype= "dashed", size=1) + #can remove this if needed
  facet_grid(OM ~ Season, labeller = labeller(Season = month_labels)) +
  scale_color_manual(values = OM_colors) +
  scale_shape_manual(values = OM_shapes) +
  theme_bw(base_size = 16) +
  labs(
    x = expression(delta^{13}*"C"),
    y = expression(delta^{15}*"N"),
    color = "OM",
    shape = "OM"
  ) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```

# Carbon (C) and Nitrogen (N)

# Fixed-Season GLMM

# SC slopes vary by Month × OM

# All slopes in tables, only significant slopes in figures

1.  Read data

```{r}
df <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_CN_YEAR_1_ALL.csv", check.names = FALSE)
```

2.  Data preparation

```{r}
df <- df %>%
  mutate(
    Site  = factor(Site),
    month = factor(Season, levels = c("February", "May", "August", "October")),
    OM    = factor(OM),
    SC    = as.numeric(SC),
    C     = as.numeric(C),
    N     = as.numeric(N)
  )
```

3.  Salinity prediction sequence

```{r}
SC_seq <- seq(
  min(df$SC, na.rm = TRUE),
  max(df$SC, na.rm = TRUE),
  length.out = 100
)
```

4.  Plot aesthetics

```{r}
OM_shapes <- c("CBOM" = 16, "FBOM" = 17, "Periphyton" = 15)
OM_colors <- c(
  "CBOM" = "#E69F00",
  "FBOM" = "#0072B2",
  "Periphyton" = "#228B22"
)

month_labels <- c(
  "February" = "February (Winter 2024)",
  "May"      = "May (Spring 2024)",
  "August"   = "August (Summer 2024)",
  "October"  = "October (Autumn 2024)"
)
```

## CARBON (C)

5.  Fit GLMM

```{r}
mod_C <- lmer(
  C ~ SC * month * OM + (1 | Site),
  data = df
)
```

6.  Extract SC slopes (table)

```{r}
C_SC_slopes <- emtrends(
  mod_C,
  ~ month | OM,
  var = "SC"
) %>%
  summary(infer = TRUE) %>%
  as.data.frame() %>%
  mutate(
    t_value = SC.trend / SE,
    p_value = 2 * pt(-abs(t_value), df),
    Significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ "ns"
    )
  ) %>%
  select(OM, month, SC.trend, SE, df,
         t_value, p_value, lower.CL, upper.CL, Significance)

print("=== Carbon (C) ~ SC Slopes by Month × OM ===")
print(C_SC_slopes)
write.csv(C_SC_slopes, "~/Downloads/Updated code + summary sheets/C_SC_slopes_by_month_OM.csv", row.names = FALSE)
```

7.  Predicted values

```{r}
emm_C <- emmeans(
  mod_C,
  ~ SC | month + OM,
  at = list(SC = SC_seq)
)

pred_C <- as.data.frame(emm_C)
```

8.  Keep only significant slopes

```{r}
sig_C <- C_SC_slopes %>%
  filter(Significance != "ns") %>%
  select(OM, month)

pred_C_sig <- pred_C %>%
  semi_join(sig_C, by = c("OM", "month"))
```

9.  Plot Carbon

```{r}
p_C <- ggplot(df, aes(x = SC, y = C, color = OM, shape = OM)) +
  geom_jitter(size = 3, width = 0.1, alpha = 0.7) +
  geom_ribbon(
    data = pred_C_sig,
    aes(x = SC, ymin = lower.CL, ymax = upper.CL,
        fill = OM, group = interaction(OM, month)),
    alpha = 0.3,
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  geom_line(
    data = pred_C_sig,
    aes(x = SC, y = emmean,
        color = OM, group = interaction(OM, month)),
    size = 1.2,
    inherit.aes = FALSE
  ) +
  scale_color_manual(values = OM_colors) +
  scale_fill_manual(values = OM_colors) +
  scale_shape_manual(values = OM_shapes) +
  facet_grid(
    OM ~ month,
    labeller = labeller(month = month_labels),
    scales = "free_y"
  ) +
  theme_bw(base_size = 16) +
  labs(
    x = "Specific Conductivity (µS/m)",
    y = "% Carbon"
  ) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 14, face = "bold")
  )

print(p_C)
```

# NITROGEN (N)

10. Fit GLMM

```{r}
mod_N <- lmer(
  N ~ SC * month * OM + (1 | Site),
  data = df
)
```

11. Extract SC slopes (table)

```{r}
N_SC_slopes <- emtrends(
  mod_N,
  ~ month | OM,
  var = "SC"
) %>%
  summary(infer = TRUE) %>%
  as.data.frame() %>%
  mutate(
    t_value = SC.trend / SE,
    p_value = 2 * pt(-abs(t_value), df),
    Significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ "ns"
    )
  ) %>%
  select(OM, month, SC.trend, SE, df,
         t_value, p_value, lower.CL, upper.CL, Significance)

print("=== Nitrogen (N) ~ SC Slopes by Month × OM ===")
print(N_SC_slopes)
write.csv(N_SC_slopes, "~/Downloads/Updated code + summary sheets/N_SC_slopes_by_month_OM.csv", row.names = FALSE)
```

12. Predicted values

```{r}
emm_N <- emmeans(
  mod_N,
  ~ SC | month + OM,
  at = list(SC = SC_seq)
)

pred_N <- as.data.frame(emm_N)
```

13. Keep only significant slopes

```{r}
sig_N <- N_SC_slopes %>%
  filter(Significance != "ns") %>%
  select(OM, month)

pred_N_sig <- pred_N %>%
  semi_join(sig_N, by = c("OM", "month"))
```

14. Plot Nitrogen

```{r}
p_N <- ggplot(df, aes(x = SC, y = N, color = OM, shape = OM)) +
  geom_jitter(size = 3, width = 0.1, alpha = 0.7) +
  geom_ribbon(
    data = pred_N_sig,
    aes(x = SC, ymin = lower.CL, ymax = upper.CL,
        fill = OM, group = interaction(OM, month)),
    alpha = 0.3,
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  geom_line(
    data = pred_N_sig,
    aes(x = SC, y = emmean,
        color = OM, group = interaction(OM, month)),
    size = 1.2,
    inherit.aes = FALSE
  ) +
  scale_color_manual(values = OM_colors) +
  scale_fill_manual(values = OM_colors) +
  scale_shape_manual(values = OM_shapes) +
  facet_grid(
    OM ~ month,
    labeller = labeller(month = month_labels),
    scales = "free_y"
  ) +
  theme_bw(base_size = 16) +
  labs(
    x = "Specific Conductivity (µS/m)",
    y = "% Nitrogen"
  ) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 14, face = "bold")
  )

print(p_N)
```

# Summary Statistics for Figures in Manuscript - AA

1.  Helper function

```{r}
summarise_var <- function(df, group_vars, value_var) {
  df %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(
      n      = sum(!is.na(.data[[value_var]])),
      mean   = round(mean(.data[[value_var]], na.rm = TRUE), 3),
      sd     = round(sd(.data[[value_var]],   na.rm = TRUE), 3),
      median = round(median(.data[[value_var]], na.rm = TRUE), 3),
      min    = round(min(.data[[value_var]],  na.rm = TRUE), 3),
      max    = round(max(.data[[value_var]],  na.rm = TRUE), 3),
      .groups = "drop"
    )
}
```

2.  Figure 1: AFDM - dataset: SALTY_C_AFDM_YEAR_1_ALL.csv

```{r}
cat("\n\n══════ AFDM Summary Statistics (Fig 1) ══════\n")

df_afdm <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_AFDM_YEAR_1_ALL.csv", check.names = FALSE) %>%
  mutate(
    Site  = factor(Site),
    OM    = factor(OM),
    month = factor(Season, levels = c("February", "May", "August", "October")),
    SC    = as.numeric(SC),
    AFDM  = as.numeric(AFDM)
  )

afdm_stats <- summarise_var(df_afdm, c("OM", "month"), "AFDM")
print(afdm_stats)
write.csv(afdm_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_AFDM_by_OM_Month.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_AFDM_by_OM_Month.csv\n")
```

3.  Figure 2: CHL-A - dataset: SALTY_C_CHLA_ROCKS.csv

```{r}
cat("\n\n══════ Chlorophyll-a Summary Statistics (Fig 2) ══════\n")

df_chla <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_CHLA_ROCKS.csv", check.names = FALSE) %>%
  mutate(
    Site  = factor(Site),
    month = factor(Season, levels = c("February", "May", "August", "October")),
    SC    = as.numeric(SC),
    CHLA  = as.numeric(CHLA)
  )

chla_stats <- summarise_var(df_chla, "month", "CHLA")
print(chla_stats)
write.csv(chla_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_CHLA_by_Month.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_CHLA_by_Month.csv\n")
```

4.  Figure 3-4: Stable Isotopes - dataset: SALTY_C_SI_YEAR_1_ALL.csv

```{r}
cat("\n\n══════ δ13C Summary Statistics (Fig 3) ══════\n")

df_si <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_SI_YEAR_1_ALL.csv", check.names = FALSE) %>%
  mutate(
    Season = factor(Season, levels = c("February", "May", "August", "October")),
    Site   = factor(Site),
    OM     = factor(OM),
    SC     = as.numeric(SC),
    d13C   = as.numeric(d13C),
    d15N   = as.numeric(d15N)
  )

d13C_stats <- summarise_var(df_si, c("OM", "Season"), "d13C")
print(d13C_stats)
write.csv(d13C_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_d13C_by_OM_Season.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_d13C_by_OM_Season.csv\n")

cat("\n\n══════ δ15N Summary Statistics (Fig 4) ══════\n")

d15N_stats <- summarise_var(df_si, c("OM", "Season"), "d15N")
print(d15N_stats)
write.csv(d15N_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_d15N_by_OM_Season.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_d15N_by_OM_Season.csv\n")

```

5.  5-7 C:N, %C, %N - dataset: SALTY_C_CN_YEAR_1_ALL.csv

```{r}
cat("\n\n══════ C:N Summary Statistics (Fig 6 – Erin flagged) ══════\n")

df_cn <- read.csv("~/Downloads/Updated code + summary sheets/SALTY_C_CN_YEAR_1_ALL.csv", check.names = FALSE) %>%
  mutate(
    Site  = factor(Site),
    month = factor(Season, levels = c("February", "May", "August", "October")),
    OM    = factor(OM),
    SC    = as.numeric(SC),
    C_N   = as.numeric(C_N),
    C     = as.numeric(C),
    N     = as.numeric(N)
  )

cn_stats <- summarise_var(df_cn, c("OM", "month"), "C_N")
print(cn_stats)
write.csv(cn_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_CN_by_OM_Month.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_CN_by_OM_Month.csv\n")
cat("  NOTE for Erin's comment: 'lower.CL / upper.CL' ribbons in Fig 6 are\n")
cat("  95% confidence intervals from the GLMM (emmeans), NOT raw ranges.\n")
cat("  Median and range are now available in the CSV above.\n")

cat("\n\n══════ %~/Downloads/Updated code + summary sheets/Carbon Summary Statistics (Fig 5) ══════\n")

c_stats <- summarise_var(df_cn, c("OM", "month"), "C")
print(c_stats)
write.csv(c_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_C_by_OM_Month.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_C_by_OM_Month.csv\n")

cat("\n\n══════ %Nitrogen Summary Statistics (Fig 7) ══════\n")

n_stats <- summarise_var(df_cn, c("OM", "month"), "N")
print(n_stats)
write.csv(n_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_N_by_OM_Month.csv", row.names = FALSE)
cat("  → Saved: ~/Downloads/Updated code + summary sheets/SummaryStats_N_by_OM_Month.csv\n")
```

6.  Combine all tables into one master summary

```{r}
all_stats <- bind_rows(
  afdm_stats %>% mutate(Figure = "Fig_AFDM", Variable = "AFDM"),
  chla_stats %>% mutate(Figure = "Fig_CHLA", Variable = "CHLA", OM = "Periphyton") %>%
    rename(month = month),
  d13C_stats %>% mutate(Figure = "Fig_d13C", Variable = "d13C") %>% rename(month = Season),
  d15N_stats %>% mutate(Figure = "Fig_d15N", Variable = "d15N") %>% rename(month = Season),
  cn_stats   %>% mutate(Figure = "Fig6_CN",  Variable = "C:N"),
  c_stats    %>% mutate(Figure = "Fig_C",    Variable = "C_pct"),
  n_stats    %>% mutate(Figure = "Fig_N",    Variable = "N_pct")
) %>%
  select(Figure, Variable, everything())

write.csv(all_stats, "~/Downloads/Updated code + summary sheets/SummaryStats_ALL_FIGURES.csv", row.names = FALSE)
cat("\n\n══════ Master file saved: ~/Downloads/Updated code + summary sheets/SummaryStats_ALL_FIGURES.csv ══════\n")
cat("Done.\n")
```
